---
title: 'Hoxell-enabled Personalization: Berna Hotel'
author: "Gabriele Piccoli"
output:
    pdf_document:
        includes:
            in_header: header.tex
---

#Introduction
This is the current report for the persalization study of *HGRM* use at the *Berna Hotel*. 
The report was generated on `r format(Sys.time(), "%d %B, %Y")`.

```{r echo=FALSE}
fileH1i <- "H1individual.csv"
fileH1a <- "H1aggregate.csv"
fileCompl <- "BernaComplete_2011_20150630.csv"
fileRep <- "BernaRepeatedPair_2011_20150630_BothChannels.csv"
fileDisinterm <- "BernaAllReservation_2011_20150630.csv"
fileRepMult <- "BernaAllRez_2011_20150630.csv"

individual <- read.csv(fileH1i)
date_range <- with(individual, range(as.Date(Date)))

Levels <- levels(individual$Date)

## OLD CODE FOR MANIPULATING DATES
# ind <- grep("^20", Levels)
# Levels[ind] <- sapply(strsplit(Levels[ind], "/"), function(x) paste(x[c(2, 3, 1)], collapse = "/"))
# levels(individual$Date) <- Levels
# individual$Month <- format(strptime(individual$Date, "%m/%d/%Y"), "%Y-%m")
##

individual$Month <- format(strptime(individual$Date, "%Y-%m-%d"), "%Y-%m")

## Hypothesis 1
aggregate <- read.csv(fileH1a)
X <- data.frame(Month = rep(aggregate$Month, 2),
                Stays = rep(aggregate$Stays, 2),
                Persons = c(aggregate$phPers, aggregate$vPers),
                count = c(aggregate$countPhPers, aggregate$countVPers),
                virtualization = !is.na(aggregate$vPers),
                method = rep(c("physical", "virtual"), each = nrow(aggregate)))
X <- na.omit(X)
```

#Basic statistics
The dataset includes **`r nrow(individual)`** reservations for **`r sum(aggregate$Stays)`** hotel stays during the **`r format(date_range[1], "%d %B, %Y")`** to **`r format(date_range[2], "%d %B, %Y")`** timeframe. Of these, **`r sum(table(individual$Method)[c("physical", "virtual")])`** are personalized either physically or through myPage.
  
##Personalization Request via Traditional Channel and myPage
```{r echo=FALSE, results='hold'}
## Table 2
tab2 <- with(aggregate, rbind(Stays = c(mean = mean(Stays), sd = sd(Stays), min = min(Stays), max = max(Stays)),
              t(sapply(c("physical", "virtual"), function(method) {
                  x <- table(factor(subset(individual, Method == method, Month)$Month))
                  c(mean = mean(x), sd = sd(x), min = min(x), max = max(x))
              })),
              do.call("rbind", with(subset(individual, Method %in% c("virtual", "physical")),
                                    tapply(P_ItemCount, factor(Method), function(x)
                                           c(mean = mean(x), sd = sd(x), min = min(x), max = max(x)))))))
tab2[,1:2] <- round(tab2[,1:2], digits = 2)
knitr::kable(tab2, digits = 2)
```


\pagebreak

#The effect of Hoxell on preference elicitation: Customers
```{r echo=FALSE, results='hide'}
g1 <- glm(P_ItemCount ~ Method, data = subset(individual, Method %in% c("virtual", "physical")),
          family = "poisson")
g2 <- glm(cbind(Persons, Stays - Persons) ~ method + virtualization, data = X, family = "binomial")
round(predict(g2, newdata = data.frame(method = c("physical", "physical", "virtual"), virtualization = c(FALSE, TRUE, TRUE)), type = "response"),
      digits = 4)
```

Our analysis shows that the introduction of Hoxell has a significant impact on the number of guests who personalize their stay at Berna Hotel. It also indicates that the introduciton of personalization featurs in Hoxell has a positive effect on personalization via the tradition (physical) method.

##Poisson and Binomial Regression Results
```{r echo=FALSE, results='hold'}
tab3 <- round(rbind(coef(summary(g1)),
                    coef(summary(g2))), digits = 2)
tab3 <- cbind(tab3, c(NA, NA, round(exp(coef(g2)), digits = 2)))
knitr::kable(tab3, digits = 2)
```
  
  --
  
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=15, fig.align='center'}
##Graphs
library(zoo)

Z  <- X

# Plot adopters and percentage fo stays 
Z$Month  <- as.character(X$Month)

## Code to fix date Errors in H1aggregate.csv
## It should no longer be needed.
# ind  <- grep("^ 20", Z$Month)
# d  <- substr(as.character(zoo::as.yearmon(Z$Month[ind])), 1, 3)
# Z$Month[ind] <- paste0(d, "-", c(rep(10, 12), 11))
# End of correction code

Z$Month  <- as.Date(paste0(Z$Month, "-01"), format = "%y-%b-%d")
Z$Month <- as.yearmon(Z$Month)


xrange <- range(Z$Month)
yrange <- range(Z$Persons)
Z$pcStays <- with(Z, Persons/Stays)
zrange <- range(Z$pcStays * 100 * 2)
#pdf(file = "charts/chart1.pdf", width = 15, height = 10)
plot(xrange, yrange, type="n", xlab="Time (Months)", ylab="Stays") 
colors <- c(rgb(1,0,0), rgb(0,0,1)) 
linetype <- 1 
plotchar <- seq(18,18+2,1)
ph <- subset(Z, method == "physical")
vp <- subset(Z, method == "virtual") 
lines(ph$Month, ph$Persons, type="b", lwd=3,
      lty=linetype[1], col=rgb(1,0,0), pch=plotchar[1]) 
lines(vp$Month, vp$Persons, type="b", lwd=3,
      lty=linetype[1], col=rgb(0,0,1), pch=plotchar[2]) 
title("Number of Personalized Stays")
legend(xrange[1], yrange[2], c("Physical", "Virtual"), cex=0.8, col=colors,lwd=3,
       pch=plotchar, lty=linetype, title="Personalization")
par(new = TRUE)
plot(xrange, zrange, type="n", axes = FALSE, xlab = "", ylab = "")
axis(side=4, at = pretty(zrange), tick = TRUE)
mtext("Percentage of Stays", side=4, line=3)
lines(vp$Month, (vp$pcStays*100), type="h", lwd=15,
      lty=linetype[1], col=rgb(0,0,1,0.2)) 
lines(ph$Month, (ph$pcStays*100), type="h", lwd=15,
      lty=linetype[1], col=rgb(1,0,0)) 
#dev.off()
```

<!--
\begin{landscape}
\includegraphics{charts/chart1}
\end{landscape}
-->

\pagebreak

#The effect of Hoxell on preference elicitation: Personalizations
This analysis focuses on the number of itmes requested by customers who personalize. 
We find a strong impact of Hoxell on the number of items requested by customers. 
Specifically when personalizing through traditional means customers chose an average of `r tab2[4, 1]` items. Conversely, when using Hoxell they select `r tab2[5, 1]` items.
  
  --
  
```{r echo=FALSE, fig.height=10, fig.width=15, fig.align='center'}
# Plot number of items 
xrange <- range(Z$Month) 
yrange <- range(Z$count)
#pdf(file = "charts/chart2.pdf", width = 15, height = 10)
plot(xrange, yrange, type="n", xlab="Time (Months)", ylab="Itmes") 
lines(ph$Month, ph$count, type="b", lwd=3,
      lty=linetype[1], col=rgb(1,0,0), pch=plotchar[1]) 
lines(vp$Month, vp$count, type="b", lwd=3,
      lty=linetype[1], col=rgb(0,0,1), pch=plotchar[2]) 
title("Number of Personalizations per Stay")
legend(xrange[1], yrange[2], c("Physical", "Virtual"), cex=0.8, col=colors,lwd=3,
       pch=plotchar, lty=linetype, title="Personalization")
#dev.off()
```

<!--
\begin{landscape}
\includegraphics{charts/chart2}

\end{landscape}
-->

\pagebreak

#Analysis of emails
Ths section provides a monthly analysis of emails sent via the Mandrill application.

```{r echo=FALSE, fig.height=10, fig.width=15, fig.align='center'}
##Extra analysis of Mandrill data

Y <- aggregate[, 1:8]

# Plot adopters and percentage fo stays 
Y$Month  <- as.character(Y$Month)
Y$Month  <- as.Date(paste0(Y$Month, "-01"), format = "%y-%b-%d")
Y$Month <- as.yearmon(Y$Month)
Y <- na.omit(Y)


Y[which(Y$Month > "Jun 2014"), c("eSent")] <- round(Y[which(Y$Month > "Jun 2014"), c("eSent")]/2, 0)
Y[which(Y$Month > "Jun 2014"), c("eDelivered")] <- round(Y[which(Y$Month > "Jun 2014"), c("eDelivered")]/2, 0)

colors <- c(rgb(1,0,0), rgb(0,0,1), rgb(0,1,0), rgb(1,0,1), rgb(0,0,0)) 
linetype <- 1 
plotchar <- seq(18,18+5,1)

# Plot emails 
xrange <- range(Y$Month) 
yrange <- range(with(Y, c(0, Stays, eSent)))
#pdf(file = "charts/chart3.pdf", width = 15, height = 10)
plot(xrange, yrange, type="n", xlab="Time (Months)", ylab="Responses") 
lines(Y$Month, Y$Stays, type="b", lwd=3,
      lty=linetype[1], col=colors[1], pch=plotchar[1]) 
lines(Y$Month, Y$eSent, type="b", lwd=3,
      lty=linetype[1], col=colors[2], pch=plotchar[2]) 
lines(Y$Month, Y$eDelivered, type="b", lwd=3,
      lty=linetype[1], col=colors[3], pch=plotchar[3])
lines(Y$Month, Y$eOpened, type="b", lwd=3,
      lty=linetype[1], col=colors[4], pch=plotchar[4])
lines(Y$Month, Y$eClicked, type="b", lwd=3,
      lty=linetype[1], col=colors[5], pch=plotchar[5])
title("Analysis of Emails: Number of messages and actions")
legend(xrange[1], yrange[2], names(Y)[2:6], cex=0.8, col=colors[1:5],lwd=3,
       pch=plotchar[1:5], lty=linetype[1], title="Totals")
#dev.off()

Y <- cbind(Y, lapply(Y[, 7:8], function(x) x/Y[, 2]))
names(Y) <- c(names(Y)[1:8], "phStays", "vStays")
Y <- cbind(Y, lapply(Y[, 7:8], function(x) x/Y[, 4]))
names(Y) <- c(names(Y)[1:10], "phDelivered", "vDelivered")
Y <- cbind(Y, lapply(Y[, 7:8], function(x) x/Y[, 5]))
names(Y) <- c(names(Y)[1:12], "phOpened", "vOpened")
Y <- cbind(Y, lapply(Y[, 7:8], function(x) x/Y[, 6]))
names(Y) <- c(names(Y)[1:14], "phClicked", "vClicked")
Y <- cbind(Y[, 1:8], round(Y[, 9:16]*100, 2))

# Plot percentages 
xrange <- range(Y$Month) 
yrange <- range(with(Y, c(0, vClicked)))
#pdf(file = "charts/chart4.pdf", width = 15, height = 10)
plot(xrange, yrange, type="n", xlab="Time (Months)", ylab="Percentage of responses") 
lines(Y$Month, Y$vStays, type="b", lwd=3,
      lty=linetype[1], col=colors[1], pch=plotchar[1]) 
lines(Y$Month, Y$vDelivered, type="b", lwd=3,
      lty=linetype[1], col=colors[2], pch=plotchar[2]) 
lines(Y$Month, Y$vOpened, type="b", lwd=3,
      lty=linetype[1], col=colors[3], pch=plotchar[3])
lines(Y$Month, Y$vClicked, type="b", lwd=3,
      lty=linetype[1], col=colors[4], pch=plotchar[4])
lines(Y$Month, Y$phStays, type="b", lwd=3,
      lty=linetype[1], col=colors[5], pch=plotchar[5])
title("Analysis of Emails: Percentage of Personalizations")
legend(x = "topright", names(Y)[c(10,12,14,16,9)], cex=0.8, col=colors[1:5],lwd=3,
       pch=plotchar[1:5], lty=linetype[1], title="Totals")
#dev.off()

##############
```

<!--
\begin{landscape}
\includegraphics{charts/chart3}

\pagebreak
\includegraphics{charts/chart4}

\end{landscape}
-->

\pagebreak

#The Effect of Personalization on Service and Value Hotel Ratings
```{r echo=FALSE, results='hide'}
## Table 4
x <- read.csv(fileCompl, encoding = "UTF-8", quote = "\"", sep = ",",
              strip.white = TRUE, as.is = TRUE) 
## OLD CODE FOR MANIPULATING DATES
# x$Departure[grep("201$", x$Departure)] <- 
#   sapply(grep("201$", x$Departure), function(w)   
#    gsub("201$", format(as.Date(x$Arrival[w], "%A, %B %d, %Y"), "%Y"), x$Departure[w]))
# x$ar <- as.Date(x$Arrival, "%A, %B %d, %Y")
# x$ar[is.na(x$ar)] <- as.Date(x$Arrival[is.na(x$ar)], "%Y/%m/%d")
# x$dp <- as.Date(x$Departure, "%A, %B %d, %Y")
# x$dp[is.na(x$dp)] <- as.Date(x$Departure[is.na(x$dp)], "%Y/%m/%d")
# x$LoS <- with(x, dp-ar)

## NEW CODE (IRIS DROPPED Departure and Arrival columns)
x$ar <- as.Date(x$ARRIVALDATE, "%Y-%m-%d")
x$dp <- as.Date(x$DEPARTUREDATE, "%Y-%m-%d")
x$LoS <- with(x, dp-ar)

is.na(x$ADR[which(x$ADR == 0)]) <- TRUE
x$ROOM_TYPE <- factor(x$ROOM_TYPE)
levels(x$ROOM_TYPE) <- gsub("^ ", "", levels(x$ROOM_TYPE))
## ROOM_TYPE = "PX" is set to NA
levs <- sapply(c("^COM", "^Q", "^ST", "^S(3|P)"), function(g) {
    u <- rep(FALSE, nlevels(x$ROOM_TYPE))
    u[grep(g, levels(x$ROOM_TYPE))] <- TRUE
    u
    })
levels(x$ROOM_TYPE) <- c(NA, "COMFORT", "QUALITY", "SINGOLA (BASIC)", "SUPERIOR")[levs %*% 1:4 + 1]
x$PP <- x$PrefPrice
x$PrefPrice[is.na(x$PrefPrice)] <- 0
is.na(x$PP[which(x$PP == 0)]) <- TRUE
dim(x)
```

We test the effect of personalizations on hotel service and value ratings using `r nrow(x)` reviews posted to Booking.com merged with reservation data.

```{r echo=FALSE, results='hide'}
tab4 <- list(Rating = t(apply(x[, c("Service", "Value")], 2, table)),
             Room = table(x$ROOM_TYPE),
             VAE = table(x$Personalization),
             round(t(sapply(x[c("ADR", "LoS", "PP", "ADULTS", "CHILDREN")], function(x) {
                 x <- as.numeric(x)
                 c(mean = mean(x, na.rm = TRUE), sd = sd(x, na.rm = TRUE), min = min(x, na.rm = TRUE), max = max(x, na.rm = TRUE))
             })), digits = 2))
tab4
x$ADR <- scale(x$ADR)
```

##Personalization Effect on Service
```{r echo=FALSE, results='hold'}
## Table 5
M <- na.omit(x[, c("Service", "ROOM_TYPE", "LoS", "ADR", "Personalization", "ADULTS", "CHILDREN", "PrefPrice")])
M$Service <- ordered(M$Service, c(2.5, 5, 7.5, 10))
m0 <- ordinal::clm(Service ~ Personalization + ADR + LoS + ROOM_TYPE + ADULTS + CHILDREN + PrefPrice, data = subset(M, ADULTS + CHILDREN > 0 & PrefPrice < 50), Hess = TRUE)
tab5 <- round(cbind(coef(summary(m0)), exp(coef(m0))), digits = 3)
tab5[1:3, ncol(tab5)] <- NA
knitr::kable(tab5, digits = 2)
```

##Personalization Effect on Value
```{r echo=FALSE, results='hold'}
## Table 6
M <- na.omit(x[, c("Value", "ROOM_TYPE", "LoS", "ADR", "Personalization", "ADULTS", "CHILDREN", "PrefPrice")])
M$Value <- ordered(M$Value, c(2.5, 5, 7.5, 10))
m0 <- ordinal::clm(Value ~ Personalization + ADR + LoS + ROOM_TYPE + ADULTS + CHILDREN + PrefPrice, data = subset(M, ADULTS + CHILDREN > 0 & PrefPrice < 50), Hess = TRUE)
tab6 <- round(cbind(coef(summary(m0)), exp(coef(m0))), digits = 3)
tab6[1:3, ncol(tab6)] <- NA
knitr::kable(tab6, digits = 2)
```

\pagebreak

##The Effect of Number of Personalization on Service
```{r echo=FALSE, results='hold'}
## EXTRA ANALYSIS OF NUMBER OF PERSONALIZATIONS
H <- na.omit(x[, c("Service", "Count.of.Personalization.Items.Used", "ROOM_TYPE", "LoS", "ADR", "Personalization", "ADULTS", "CHILDREN", "PrefPrice")])
H$Service <- ordered(H$Service, c(2.5, 5, 7.5, 10))
h0 <- ordinal::clm(Service ~ Count.of.Personalization.Items.Used + ADR + LoS + ROOM_TYPE + ADULTS + CHILDREN + PrefPrice, data = subset(H, Personalization == 1 & Count.of.Personalization.Items.Used < 30 & ADULTS + CHILDREN > 0 & PrefPrice < 50), Hess = TRUE)
tabH <- round(cbind(coef(summary(h0)), exp(coef(h0))), digits = 3)
tabH[1:3, ncol(tabH)] <- NA
knitr::kable(tabH, digits = 2)
```

##The Effect of Number of Personalization on Value
```{r echo=FALSE, results='hold'}
H <- na.omit(x[, c("Value", "Count.of.Personalization.Items.Used", "ROOM_TYPE", "LoS", "ADR", "Personalization", "ADULTS", "CHILDREN", "PrefPrice")])
H$Value <- ordered(H$Value, c(2.5, 5, 7.5, 10))
h0 <- ordinal::clm(Value ~ Count.of.Personalization.Items.Used + ADR + LoS + ROOM_TYPE + ADULTS + CHILDREN + PrefPrice, data = subset(H, Personalization == 1 & Count.of.Personalization.Items.Used < 30 & ADULTS + CHILDREN > 0 & PrefPrice < 50), Hess = TRUE)
tabH <- round(cbind(coef(summary(h0)), exp(coef(h0))), digits = 3)
tabH[1:3, ncol(tabH)] <- NA
knitr::kable(tabH, digits = 2)
##
```

\pagebreak

#Hoxell Impact on Positive Share-Shift
This analysis focuses on the effect of personalization on the creation of a direct relationship with the customers that leads to share-shift from intermediated (high-cost) to direct (low-cost) distribution channels. 

```{r echo=FALSE, results='hide', warning=FALSE}
## Hypothesis 3
repeated <- read.csv(fileRep, sep = ",")
dim(repeated)
is.na(repeated$ADR[which(repeated$ADR == 0)]) <- TRUE
repeated$ROOM_TYPE <- factor(repeated$ROOM_TYPE)
levels(repeated$ROOM_TYPE) <- gsub("^ ", "", levels(repeated$ROOM_TYPE))
## ROOM_TYPE = "PM" and "PX" is set to NA
levs <- sapply(c("^COM", "^Q", "^ST", "^S(3|P)"), function(g) {
    u <- rep(FALSE, nlevels(repeated$ROOM_TYPE))
    u[grep(g, levels(repeated$ROOM_TYPE))] <- TRUE
    u
    })
levels(repeated$ROOM_TYPE) <- c(NA, "COMFORT", "QUALITY", "SINGOLA (BASIC)", "SUPERIOR")[levs %*% 1:4 + 1]
repeated$ar <- as.Date(repeated$ARRIVALDATE, "%Y-%m-%d")
repeated$dp <- as.Date(repeated$DEPARTUREDATE, "%Y-%m-%d")
repeated$LoS <- with(repeated, dp-ar)
repeated$Method <- factor(tolower(repeated$Method), c("", "physical", "virtual"), c("None", "Physical", "Virtual"))
repeated$ADR <- scale(repeated$ADR)

length(unique(repeated$GUEST_ID))
table(subset(repeated, Repeated == 0, Personalization))

channel <- repeated
internet_channels <- c("HOTEL.DE AG", "HRS-HOTEL RESERVATION SERVICE", "INITALIA.IT",
                       "Laterooms", "LATEROOMS LTD", "Orbitz", "ORBITZ", "ORBITZ WORLDWIDE INC.",
  	       "VENERE NET", "AGODA COMPANY PTE, LTD", "ALESIA DISTRIBUTION",
		       "EASYTOBOOK.COM", "HOTELREZ LTD", "LASTMINUTE.COM", "Lastminute (GHE)",
		       "RATESTOGO", "TRANSHOTEL", "TRANSHOTEL.", "TRAVEL CONSULT OY", "TRAVELEUROPE",
		       "WOTIF.COM")
gds_channels <- c("Sabre", "GALILEO", "Amadeus", "Worldspan", "Pegasus")
expedia <- c("EXPEDIA (HC) TRAVEL CORPORATED", "EXPEDIA TRAVEL CORPORATED")
hp <- c("HOME PAGE", "Website")
cc <- "Call Center Berna"
new_levels <- c("", "bookings europe b.v.", "hotel berna milano", "home page", "internet", "gds", "expedia")
levs <- sapply(c(as.list(new_levels[1:3]), list(hp), list(internet_channels), list(gds_channels), list(expedia), list(cc)), function(g) {
    u <- rep(FALSE, nlevels(channel$BOOKING_CHANNEL))
    u[tolower(levels(channel$BOOKING_CHANNEL)) %in% tolower(g)] <- TRUE
    u
    })

levels(channel$BOOKING_CHANNEL) <- c(NA, new_levels)[levs %*% seq_len(ncol(levs)) + 1]
wide <- reshape(channel[, c("Personalization", "GUEST_ID", "Repeated", "BOOKING_CHANNEL", "ADR", "ROOM_TYPE", "ADULTS", "CHILDREN", "LoS", "Method")], 
                timevar = "Repeated", idvar = "GUEST_ID", direction = "wide")

dim(wide)
table(wide$Method.0)
table(wide$BOOKING_CHANNEL.0 %in% c("bookings europe b.v.", "expedia", "internet"))
table(wide$BOOKING_CHANNEL.0 == "home page")
```

The analysis is based on a sample of `r nrow(wide)` reservations by guests who visited the hotel more than once during the timeframe of our analysis. Of these returning customers `r nrow(wide) - sum(wide$Method.0 == "None")` engaged in personalization during their first visit to the Berna Hotel.

##Beneficial Share-Shift
```{r echo=FALSE, results='hold'}
## Table 7
g1 <- glm(BOOKING_CHANNEL.1 == "home page" ~ I(Method.0 == "Virtual") + ADR.1 + ROOM_TYPE.1 + ADULTS.1 + CHILDREN.1, 
          data = subset(wide, !BOOKING_CHANNEL.0 %in% c("", "gds", "hotel berna milano", "home page")), family = "binomial")
knitr::kable(round(cbind(coef(summary(g1)), Odds.Ratio = exp(coef(g1))), digits = 3))

agg <- with(subset(wide, !BOOKING_CHANNEL.0 %in% c("", "gds", "hotel berna milano", "home page")),
            table(Virtual = Method.0 == "Virtual", Homepage = BOOKING_CHANNEL.1 == "home page"))
prop.table(agg, 1)[2,2] / prop.table(agg, 1)[1,2] - 1
```

##Detrimental Share-Shift
```{r echo=FALSE, results='hold'}
## Table 8
g2 <- glm(BOOKING_CHANNEL.1 != "home page" ~ I(Method.0 == "Virtual") + ADR.1 + ROOM_TYPE.1 + ADULTS.1 + CHILDREN.1, 
          data = subset(wide, BOOKING_CHANNEL.0 == "home page"), family = "binomial")
knitr::kable(round(cbind(coef(summary(g2)), Odds.Ratio = exp(coef(g2))), digits = 3)) 

agg <- with(subset(wide, BOOKING_CHANNEL.0  == "home page"),
            table(Virtual = Method.0 == "Virtual", Homepage = BOOKING_CHANNEL.1 != "home page"))
prop.table(agg, 1)[2,2] / prop.table(agg, 1)[1,2] - 1
```


\pagebreak

#Hoxell Impact on Positive Share-Shift - Limited window
This analysis focuses on the effect of personalization on the creation of a direct relationship with the customers that leads to share-shift from intermediated (high-cost) to direct (low-cost) distribution channels.  Unlike the previous analysis we only use a portion of the data constraining only to people who return to the property within a limited window of days

```{r echo=FALSE, results='hide', warning=FALSE}
#Hypothesis 3 - with limited number of days for repeated window
channel <- repeated

mf <- merge(channel[channel$NoRepeat == 1, c("GUEST_ID", "DEPARTUREDATE")], 
            channel[channel$NoRepeat == 2, c("GUEST_ID", "ARRIVALDATE",  "BOOKING_DATE")])
mf$DEPARTUREDATE <- as.Date(mf$DEPARTUREDATE, format = "%m/%d/%y")
mf$ARRIVALDATE <- as.Date(mf$ARRIVALDATE, format = "%m/%d/%y")
mf$BOOKING_DATE <- as.Date(mf$BOOKING_DATE, format ="%m/%d/%y")
mf$dif <- with(mf, ARRIVALDATE - DEPARTUREDATE)
mf$dif2 <- with(mf, BOOKING_DATE - DEPARTUREDATE)
wnd <- 365
mf <- mf[which(mf$dif > 0 & mf$dif < wnd), ]
mf <- mf[which(mf$dif2 > 0), ]
channel <- channel[channel$GUEST_ID %in% mf$GUEST_ID, ]

internet_channels <- c("HOTEL.DE AG", "HRS-HOTEL RESERVATION SERVICE", "INITALIA.IT",
                       "Laterooms", "LATEROOMS LTD", "Orbitz", "ORBITZ", "ORBITZ WORLDWIDE INC.",
                       "VENERE NET", "AGODA COMPANY PTE, LTD", "ALESIA DISTRIBUTION",
                       "EASYTOBOOK.COM", "HOTELREZ LTD", "LASTMINUTE.COM", "Lastminute (GHE)",
                       "RATESTOGO", "TRANSHOTEL", "TRANSHOTEL.", "TRAVEL CONSULT OY", "TRAVELEUROPE",
                       "WOTIF.COM")
gds_channels <- c("Sabre", "GALILEO", "Amadeus", "Worldspan", "Pegasus")
expedia <- c("EXPEDIA (HC) TRAVEL CORPORATED", "EXPEDIA TRAVEL CORPORATED")
hp <- c("HOME PAGE", "Website")
cc <- "Call Center Berna"
new_levels <- c("", "bookings europe b.v.", "hotel berna milano", "home page", "internet", "gds", "expedia")
levs <- sapply(c(as.list(new_levels[1:3]), list(hp), list(internet_channels), list(gds_channels), list(expedia), list(cc)), function(g) {
  u <- rep(FALSE, nlevels(channel$BOOKING_CHANNEL))
  u[tolower(levels(channel$BOOKING_CHANNEL)) %in% tolower(g)] <- TRUE
  u
})

levels(channel$BOOKING_CHANNEL) <- c(NA, new_levels)[levs %*% seq_len(ncol(levs)) + 1]
wide <- reshape(channel[, c("Personalization", "GUEST_ID", "Repeated", "BOOKING_CHANNEL", "ADR", "ROOM_TYPE", "ADULTS", "CHILDREN", "LoS", "Method")], 
                timevar = "Repeated", idvar = "GUEST_ID", direction = "wide")

dim(wide)
table(wide$Method.0)
table(wide$BOOKING_CHANNEL.0 %in% c("bookings europe b.v.", "expedia", "internet"))
table(wide$BOOKING_CHANNEL.0 == "home page")
```

The analysis is based on a sample of `r nrow(wide)` reservations by guests who visited the hotel more than once during the timeframe of our analysis. However we limit the analysis to only guests who returned withing **`r wnd`** days. Of these returning customers `r nrow(wide) - sum(wide$Method.0 == "None")` engaged in personalization during their first visit to the Berna Hotel. 
This analysis is performed by focusing only on those second stays that were booked after departure of the first stay. In other words, we eliminated instances where the second reservation was made before the first stay as the guest would have booked before enjoying the benefits of personalization.

##Beneficial Share-Shift
```{r echo=FALSE, results='hold'}
## Table 7b
g1 <- glm(BOOKING_CHANNEL.1 == "home page" ~ I(Method.0 == "Virtual") + ADR.1 + ROOM_TYPE.1 + ADULTS.1 + CHILDREN.1, 
          data = subset(wide, !BOOKING_CHANNEL.0 %in% c("", "gds", "hotel berna milano", "home page")), family = "binomial")
knitr::kable(round(cbind(coef(summary(g1)), Odds.Ratio = exp(coef(g1))), digits = 3))

agg <- with(subset(wide, !BOOKING_CHANNEL.0 %in% c("", "gds", "hotel berna milano", "home page")),
            table(Virtual = Method.0 == "Virtual", Homepage = BOOKING_CHANNEL.1 == "home page"))
prop.table(agg, 1)[2,2] / prop.table(agg, 1)[1,2] - 1
```

##Detrimental Share-Shift
```{r echo=FALSE, results='hold'}
## Table 8b
g2 <- glm(BOOKING_CHANNEL.1 != "home page" ~ I(Method.0 == "Virtual") + ADR.1 + ROOM_TYPE.1 + ADULTS.1 + CHILDREN.1, 
          data = subset(wide, BOOKING_CHANNEL.0 == "home page"), family = "binomial")
knitr::kable(round(cbind(coef(summary(g2)), Odds.Ratio = exp(coef(g2))), digits = 3))

agg <- with(subset(wide, BOOKING_CHANNEL.0  == "home page"),
            table(Virtual = Method.0 == "Virtual", Homepage = BOOKING_CHANNEL.1 != "home page"))
prop.table(agg, 1)[2,2] / prop.table(agg, 1)[1,2] - 1
```

```{r echo=FALSE, results='hide', warning=FALSE}
#Hypothesis 3 - with limited number of days for repeated window
channel <- repeated

mf <- merge(channel[channel$NoRepeat == 1, c("GUEST_ID", "DEPARTUREDATE")], 
            channel[channel$NoRepeat == 2, c("GUEST_ID", "ARRIVALDATE",  "BOOKING_DATE")])
mf$DEPARTUREDATE <- as.Date(mf$DEPARTUREDATE, format = "%m/%d/%y")
mf$ARRIVALDATE <- as.Date(mf$ARRIVALDATE, format = "%m/%d/%y")
mf$BOOKING_DATE <- as.Date(mf$BOOKING_DATE, format ="%m/%d/%y")
mf$dif <- with(mf, ARRIVALDATE - DEPARTUREDATE)
mf$dif2 <- with(mf, BOOKING_DATE - DEPARTUREDATE)
wnd <- 183
mf <- mf[which(mf$dif > 0 & mf$dif < wnd), ]
mf <- mf[which(mf$dif2 > 0), ]
channel <- channel[channel$GUEST_ID %in% mf$GUEST_ID, ]

internet_channels <- c("HOTEL.DE AG", "HRS-HOTEL RESERVATION SERVICE", "INITALIA.IT",
                       "Laterooms", "LATEROOMS LTD", "Orbitz", "ORBITZ", "ORBITZ WORLDWIDE INC.",
                       "VENERE NET", "AGODA COMPANY PTE, LTD", "ALESIA DISTRIBUTION",
                       "EASYTOBOOK.COM", "HOTELREZ LTD", "LASTMINUTE.COM", "Lastminute (GHE)",
                       "RATESTOGO", "TRANSHOTEL", "TRANSHOTEL.", "TRAVEL CONSULT OY", "TRAVELEUROPE",
                       "WOTIF.COM")
gds_channels <- c("Sabre", "GALILEO", "Amadeus", "Worldspan", "Pegasus")
expedia <- c("EXPEDIA (HC) TRAVEL CORPORATED", "EXPEDIA TRAVEL CORPORATED")
hp <- c("HOME PAGE", "Website")
cc <- "Call Center Berna"
new_levels <- c("", "bookings europe b.v.", "hotel berna milano", "home page", "internet", "gds", "expedia")
levs <- sapply(c(as.list(new_levels[1:3]), list(hp), list(internet_channels), list(gds_channels), list(expedia), list(cc)), function(g) {
  u <- rep(FALSE, nlevels(channel$BOOKING_CHANNEL))
  u[tolower(levels(channel$BOOKING_CHANNEL)) %in% tolower(g)] <- TRUE
  u
})

levels(channel$BOOKING_CHANNEL) <- c(NA, new_levels)[levs %*% seq_len(ncol(levs)) + 1]
wide <- reshape(channel[, c("Personalization", "GUEST_ID", "Repeated", "BOOKING_CHANNEL", "ADR", "ROOM_TYPE", "ADULTS", "CHILDREN", "LoS", "Method")], 
                timevar = "Repeated", idvar = "GUEST_ID", direction = "wide")

dim(wide)
table(wide$Method.0)
table(wide$BOOKING_CHANNEL.0 %in% c("bookings europe b.v.", "expedia", "internet"))
table(wide$BOOKING_CHANNEL.0 == "home page")
```

The analysis is based on a sample of `r nrow(wide)` reservations by guests who visited the hotel more than once during the timeframe of our analysis. However we limit the analysis to only guests who returned withing **`r wnd`** days. Of these returning customers `r nrow(wide) - sum(wide$Method.0 == "None")` engaged in personalization during their first visit to the Berna Hotel.
This analysis is performed by focusing only on those second stays that were booked after departure of the first stay. In other words, we eliminated instances where the second reservation was made before the first stay as the guest would have booked before enjoying the benefits of personalization.

##Beneficial Share-Shift
```{r echo=FALSE, results='hold'}
## Table 7b
g1 <- glm(BOOKING_CHANNEL.1 == "home page" ~ I(Method.0 == "Virtual") + ADR.1 + ROOM_TYPE.1 + ADULTS.1 + CHILDREN.1, 
          data = subset(wide, !BOOKING_CHANNEL.0 %in% c("", "gds", "hotel berna milano", "home page")), family = "binomial")
knitr::kable(round(cbind(coef(summary(g1)), Odds.Ratio = exp(coef(g1))), digits = 3))

agg <- with(subset(wide, !BOOKING_CHANNEL.0 %in% c("", "gds", "hotel berna milano", "home page")),
            table(Virtual = Method.0 == "Virtual", Homepage = BOOKING_CHANNEL.1 == "home page"))
prop.table(agg, 1)[2,2] / prop.table(agg, 1)[1,2] - 1
```

##Detrimental Share-Shift
```{r echo=FALSE, results='hold'}
## Table 8b
g2 <- glm(BOOKING_CHANNEL.1 != "home page" ~ I(Method.0 == "Virtual") + ADR.1 + ROOM_TYPE.1 + ADULTS.1 + CHILDREN.1, 
          data = subset(wide, BOOKING_CHANNEL.0 == "home page"), family = "binomial")
knitr::kable(round(cbind(coef(summary(g2)), Odds.Ratio = exp(coef(g2))), digits = 3))

agg <- with(subset(wide, BOOKING_CHANNEL.0  == "home page"),
            table(Virtual = Method.0 == "Virtual", Homepage = BOOKING_CHANNEL.1 != "home page"))
prop.table(agg, 1)[2,2] / prop.table(agg, 1)[1,2] - 1
```



\pagebreak

#Extra analysis of personalization by number of stays
The first histogram shows how the overwealming majority of guests does not personalize their stay, or always personalizes the stay. A few guests personalize some of the times with the only notable group being the 50%.
The second chart shows the proportion of guests that personalize for each group returning between zero and 20 times to the property.  

```{r echo=FALSE}
repmult <- read.csv(fileRepMult)
q <- as.data.frame(table(repmult$GUEST_ID))
q <- q[order(q$Freq, decreasing = TRUE) ,]
names(q) <- c("GUEST_ID", "NoRepeat")
z <- as.data.frame(table(repmult$GUEST_ID, repmult$Personalization))
z <- z[z$Var2 == 1, c(1, 3)]
names(z) <- c("GUEST_ID", "NoPersonalize")
q <- merge(q, z)
q <- q[order(q$NoRepeat, decreasing = TRUE) ,]
q$perc <- with(q, round(NoPersonalize/NoRepeat*100, 2))
hist(q$perc, breaks = seq(0, 100, 1), main = "Percentage of personalization")
hist(q[q$NoRepeat>1, ]$perc, breaks = seq(0, 100, 1), main = "Percentage of personalization")
```

```{r echo=FALSE}
t <- t(with(repmult, table(Personalization, NoRepeat)))
t <- round(prop.table(t, 1), 5)
barplot(t(t[1:20, 1:2]))
```

\pagebreak

These are the proportion of personalization by year for those retunring zero to four times at Berna.

```{r echo=FALSE, results='show'}
q <- repmult[which(repmult$NoRepeat < 5) , ]
t <- table(q$Personalization, q$NoRepeat, q$Year)
par(mfrow = c(2, 3))
for (i in 1:5) {
  barplot(prop.table(t[1:2, 1:5, i], 2), 
          main = paste0("Personalization in ", dimnames(t)[[3]][i]),
          xlab = "Customer Returning Times")
}
for (i in 1:5) {
  cat(paste0("Personalization in ", dimnames(t)[[3]][i]))
  print(round(t[1:2, 1:5, i], 2), 3)
}
```

\pagebreak

The following analysis shows the extent of personalization by booking channel. Berna is most successful in stimulating personalizations via the website. Given the value of personlaization, Berna should be much more proactive in stimulating personalization - particularly from guests who book directly with the hotel. 

```{r echo=FALSE, results='show'}
levs <- sapply(c(as.list(new_levels[1:3]), list(hp), list(internet_channels), list(gds_channels), list(expedia), list(cc)), function(g) {
  u <- rep(FALSE, nlevels(repmult$BOOKING_CHANNEL))
  u[tolower(levels(repmult$BOOKING_CHANNEL)) %in% tolower(g)] <- TRUE
  u
})
levels(repmult$BOOKING_CHANNEL) <- c(NA, new_levels)[levs %*% seq_len(ncol(levs)) + 1]
with(repmult, table(Personalization, BOOKING_CHANNEL))
round(prop.table(with(repmult, table(Personalization, BOOKING_CHANNEL)), 2),2)
```

\pagebreak

#Extra analysis of Traces and Personaization Items
```{r echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(parallel)
options(mc.cores = detectCores(all.tests = FALSE, logical = FALSE))
#options(mc.cores = 1L)
library(tm)

##Analysis of Traces
#p is the pattern to be search from kw
#d is the charachter vector to search into
keywCount  <- function(d, p) {
  found <- grep(paste("\\b", p, "\\b", sep = ""), d)
  if(length(found) > 0) {
    num <- length(found)
  }
  else {
    num <- 0
  }
  return(num)
}

dfCreate  <- function(p, d){
  #m <- lapply(p, function(x) sapply(d, keywCount, x, USE.NAMES = FALSE))
  m <- mclapply(p, function(x) sapply(d, keywCount, x, USE.NAMES = FALSE))
  return(m)
}
```

##Analysis of Traces - Items Requested through Physical Personalization
```{r echo=FALSE, results='hide'}
tfile  <- "NewTracesForAmenitiesCounting_only2_clean.csv"
kwfile  <- "../data_2011_2014/PhysicalRequestKeywordsAll.csv"

kw  <- read.csv(kwfile, stringsAsFactor = FALSE, strip.white = TRUE)
kw  <- kw[, -1]
kw  <- unlist(kw)
kw  <- kw[kw != ""]
names(kw) <- gsub('\\d', "", names(kw))
kw  <- kw[!duplicated(kw)]
kw  <- iconv(kw, to = "ASCII", sub = "")
#drops  <- c("bath", "baths", "pillow", "pillows")
#kw  <- kw[!kw %in% drops]
traces  <- read.csv(tfile, stringsAsFactor = FALSE, strip.white = TRUE)
traces  <- as.character(traces[ , "internalComment"])
traces  <- iconv(traces,to = "ASCII", sub = "")

#traces  <- traces[1:1000]

df  <- as.data.frame(dfCreate(kw, traces))
colnames(df)  <- kw
df$text  <- traces
df  <- df[ , c(length(df), 1:(length(df)-1))]
l  <- split(colSums(df[, -1]), names(kw))
rez <- sapply(unique(names(kw)), function(x) sum(rowSums(df[ , which(names(kw) == x) + 1]) > 0))
l2  <- l[7:8]
rez[c("Minibar", "In.room.amenities")]  <- sapply(l2, sum)[c("Minibar", "In.room.amenities")]
sum(rez)
round(rez/sum(rez), 3)
sort(round(prop.table(rez), 3))
round(cumsum(sort(prop.table(rez), decreasing = TRUE)), 3)

#This dataframe only has reviews with at least one match
ddf  <- df[sort(unique(unlist(sapply(df[ , -1], function(x) which(x > 0))))), c("text", names(which(summary(lapply(df[ , -1], function(x) which(x > 0)))[ , 1] > 0)))]

#write.table(ddf, file = "rez.csv", sep = ",", row.names = FALSE)
#save(l, file = "obj.Rdata")
```

This analysis is performed on `r sum(rez)` reservations with traces that express personalization items.
The list of items, ordered by highest percentage of requests is:
```{r}
round(cumsum(sort(prop.table(rez), decreasing = TRUE)), 3)
```


##Analysis of Items Requested through Hoxell
```{r echo=FALSE, results='hide'}
kwfile  <-  "../data_2011_2014/BernaPrefKey_categorized.csv"
chfile <- "allPreferences.csv"
kw  <- read.csv(kwfile, stringsAsFactor = FALSE, strip.white = TRUE)
kw$Category  <- factor(kw$Category)
table(kw$Category)
ch  <- read.csv(chfile, stringsAsFactor = FALSE, strip.white = TRUE)
ch  <- ch[, -(ncol(ch)-2:0)]
ch  <- ch[-(which(ch$X11 == 11)), ]
ch[is.na(ch)]  <- 0
ch[, 4:ncol(ch)]  <- as.integer(as.logical(as.matrix(ch[ , 4:ncol(ch)])))
ch$Combined  <- gsub("'", "", ch$Combined, fixed = TRUE)
l <- strsplit(ch$Combined, "(,|-)")
l <- lapply(l, unique) ## Not needed
m  <- matrix(0L, ncol = nrow(kw), nrow = nrow(ch))
colnames(m)  <- kw$preferenceID
preferences  <- as.integer(unlist(l))
stay  <- rep(seq_along(l), sapply(l, length))
pm  <- cbind(stay, preferences)[preferences > 0, ]
pm  <- cbind(pm, as.integer(factor(pm[ , 2], levels = colnames(m))))
m[pm[ , c(1, 3)]] <- 1
pos <- which(ch[, c(4:(ncol(ch)-2),ncol(ch))] > 0, arr.ind = TRUE)
zz <- gsub("X", "", colnames(ch))[-c(1:3,41)]
pos  <- cbind(pos, as.integer(factor(zz[pos[, 2]], levels = colnames(m))))
m[pos[ , c(1, 3)]] <- 1
cs  <- colSums(m)
names(cs)  <- kw$AmenityName
l <- split(cs, kw$Category)
rez <- sapply(l, sum)
sort(round(prop.table(rez), 3))
round(cumsum(sort(prop.table(rez), decreasing = TRUE)), 3)
sum(sapply(l, sum))
sum(sapply(l, sum))/nrow(m)
```

The list of items, ordered by highest percentage of requests is:

```{r}
round(cumsum(sort(prop.table(rez), decreasing = TRUE)), 3)
```

\pagebreak

#Extra analysis of Disintermediation over time

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, fig.height=9, fig.width=15, fig.align='center'}
x <- read.csv(fileDisinterm)
table(x$BOOKING_CHANNEL)

internet_channels <- c("HOTEL.DE AG", "HRS-HOTEL RESERVATION SERVICE", "INITALIA.IT",
                       "Laterooms", "LATEROOMS LTD", "Orbitz", "ORBITZ", "ORBITZ WORLDWIDE INC.",
                       "VENERE NET", "AGODA COMPANY PTE, LTD", "ALESIA DISTRIBUTION",
                       "EASYTOBOOK.COM", "HOTELREZ LTD", "LASTMINUTE.COM", "Lastminute (GHE)",
                       "RATESTOGO", "TRANSHOTEL", "TRANSHOTEL.", "TRAVEL CONSULT OY", "TRAVELEUROPE",
                       "WOTIF.COM")
gds_channels <- c("Sabre", "GALILEO", "Amadeus", "Worldspan", "Pegasus")
expedia <- c("EXPEDIA (HC) TRAVEL CORPORATED", "EXPEDIA TRAVEL CORPORATED")
hp <- c("HOME PAGE", "Website")
cc <- "Call Center Berna"
new_levels <- c("", "bookings europe b.v.", "hotel berna milano", "home page", 
                "other internet channels", "gds", "expedia", "Call Center Berna")
levs <- sapply(c(as.list(new_levels[1:3]), list(hp), list(internet_channels), list(gds_channels), 
                 list(expedia), list(cc)), function(g) {
                   u <- rep(FALSE, nlevels(x$BOOKING_CHANNEL))
                   u[tolower(levels(x$BOOKING_CHANNEL)) %in% tolower(g)] <- TRUE
                   u
})
levels(x$BOOKING_CHANNEL) <- c(NA, new_levels)[levs %*% seq_len(ncol(levs)) + 1]
X <- x[, c(1, 13:14, 24)]
X$ARRIVALDATE <- as.Date(X$ARRIVALDATE)
X$DEPARTUREDATE <- as.Date(X$DEPARTUREDATE)

library(zoo)
X$Month <- as.yearmon(X$ARRIVALDATE)

prop_ch <- sapply(unique(X$Month), function(m) {
  f <- X[X$Month == m , "BOOKING_CHANNEL"]
  prop.table(table(f))
})
colnames(prop_ch) <- unique(X$Month)
Z <- as.data.frame(t(prop_ch))
Z$Month <- as.yearmon(as.numeric(row.names(Z)))
names(Z)

colnames(Z)[1] <- "Missing"
#Z <- Z[ , -c(1,3)]
Z <- Z[order(Z$Month) , ]

###############
##Graphs

xrange <- range(Z$Month)
yrange <- range(0, 0.6)
#pdf(file = "charts/chart5.pdf", width = 15, height = 10)
plot(xrange, yrange, type="n", xlab="Time (Months)", ylab="Percentage") 
colors <- rainbow(length(colnames(Z))-1)
linetype <- 1:(length(colnames(Z))-1)
plotchar <- seq(18,18+(length(colnames(Z))-2),1)
for (i in 1:(length(colnames(Z))-1)) {
  lines(Z$Month, Z[ , i], type="b", lwd=2,
      lty=linetype[i], col=colors[i]) 
}
title("Reservations by Channel")
for (i in 1:(length(colnames(Z))-1)) {
  legend("topleft", colnames(Z)[1:(length(colnames(Z))-1)], cex=0.7, col=colors, lwd=2,
         lty=linetype, bty = "n")
}
#dev.off()

Z$OTA <- Z$`bookings europe b.v.` + Z$expedia +Z$`other internet channels`
Z$direct <- Z$`home page` + Z$`hotel berna milano` + Z$`Call Center Berna`
Z <- Z[c(1, 3, 10, 11, 9)]


xrange <- range(Z$Month)
yrange <- range(0, max(Z[ , 1:4]))
#pdf(file = "charts/chart6.pdf", width = 15, height = 10)
plot(xrange, yrange, type="n", xlab="Time (Months)", ylab="Percentage") 
colors <- rainbow(length(colnames(Z))-1)
linetype <- 1:(length(colnames(Z))-1)
plotchar <- seq(18,18+(length(colnames(Z))-2),1)
for (i in 1:(length(colnames(Z))-1)) {
  lines(Z$Month, Z[ , i], type="b", lwd=2,
        lty=linetype[i], col=colors[i]) 
}
title("Reservations by Channel")
for (i in 1:(length(colnames(Z))-1)) {
  legend("topleft", colnames(Z)[1:(length(colnames(Z))-1)], cex=0.7, col=colors, lwd=2,
         lty=linetype, bty = "n")
}
#dev.off()
```

<!--
\begin{landscape}
\includegraphics{charts/chart5}

\pagebreak
\includegraphics{charts/chart6}

\end{landscape}
-->

```{r, echo=FALSE, results='hide'}
repeated <- read.csv(fileDisinterm)
dim(repeated)
is.na(repeated$ADR[which(repeated$ADR == 0)]) <- TRUE
repeated$ROOM_TYPE <- factor(repeated$ROOM_TYPE)
levels(repeated$ROOM_TYPE) <- gsub("^ ", "", levels(repeated$ROOM_TYPE))
## ROOM_TYPE = "PM" and "PX" is set to NA
levs <- sapply(c("^COM", "^Q", "^ST", "^S(3|P)"), function(g) {
  u <- rep(FALSE, nlevels(repeated$ROOM_TYPE))
  u[grep(g, levels(repeated$ROOM_TYPE))] <- TRUE
  u
})
levels(repeated$ROOM_TYPE) <- c(NA, "COMFORT", "QUALITY", "SINGOLA (BASIC)", "SUPERIOR")[levs %*% 1:4 + 1]
repeated$ar <- as.Date(as.character(repeated$ARRIVALDATE), "%Y-%m-%d")
repeated$dp <- as.Date(as.character(repeated$DEPARTUREDATE), "%Y-%m-%d")
repeated$bd <- as.Date(repeated$BOOKING_DATE, format ="%Y-%m-%d")
repeated$LoS <- with(repeated, dp-ar)
#repeated$ADR <- scale(repeated$ADR)

length(unique(repeated$GUEST_ID))

channel <- repeated
with(channel, table(Repeated, NoRepeat))

mf <- merge(channel[channel$NoRepeat == 1, c("GUEST_ID", "dp")], 
            channel[channel$NoRepeat == 2, c("GUEST_ID", "ar",  "bd")])
mf$dif <- with(mf, ar - dp)
mf$dif2 <- with(mf, bd - dp)
# wnd <- 365
# mf <- mf[which(mf$dif > 0 & mf$dif < wnd), ]

channel <- channel[channel$NoRepeat != 0, ]
channel <- channel[channel$NoRepeat <= 10, ]
dim(channel)
mf <- mf[!is.na(mf$dif2), ]
mf <- mf[which(!mf$dif2 <= 0), ]
channel <- channel[channel$GUEST_ID %in% mf$GUEST_ID, ]
dim(channel)
with(channel, table(Repeated, NoRepeat))

internet_channels <- c("HOTEL.DE AG", "HRS-HOTEL RESERVATION SERVICE", "INITALIA.IT",
                       "Laterooms", "LATEROOMS LTD", "Orbitz", "ORBITZ", "ORBITZ WORLDWIDE INC.",
                       "VENERE NET", "AGODA COMPANY PTE, LTD", "ALESIA DISTRIBUTION",
                       "EASYTOBOOK.COM", "HOTELREZ LTD", "LASTMINUTE.COM", "Lastminute (GHE)",
                       "RATESTOGO", "TRANSHOTEL", "TRANSHOTEL.", "TRAVEL CONSULT OY", "TRAVELEUROPE",
                       "WOTIF.COM")
gds_channels <- c("Sabre", "GALILEO", "Amadeus", "Worldspan", "Pegasus")
expedia <- c("EXPEDIA (HC) TRAVEL CORPORATED", "EXPEDIA TRAVEL CORPORATED")
hp <- c("HOME PAGE", "Website")
cc <- "Call Center Berna"
new_levels <- c("", "bookings europe b.v.", "hotel berna milano", "home page", "internet", "gds", "expedia")
levs <- sapply(c(as.list(new_levels[1:3]), list(hp), list(internet_channels), list(gds_channels), list(expedia), list(cc)), 
               function(g) {
                 u <- rep(FALSE, nlevels(channel$BOOKING_CHANNEL))
                 u[tolower(levels(channel$BOOKING_CHANNEL)) %in% tolower(g)] <- TRUE
                 u
                 }
               )

levels(channel$BOOKING_CHANNEL) <- c(NA, new_levels)[levs %*% seq_len(ncol(levs)) + 1]
wide <- reshape(channel[, c("NoRepeat", "GUEST_ID", "Personalization", "BOOKING_CHANNEL", "ADR")], 
                timevar = "NoRepeat", idvar = "GUEST_ID", direction = "wide")

dim(wide)
table(wide$Personalization.1)
table(wide$Personalization.2)
table(wide$BOOKING_CHANNEL.1 %in% c("bookings europe b.v.", "expedia", "internet"))
table(wide$BOOKING_CHANNEL.1 == "home page")
```

##Stickiness of the personalization effect
The following data shows that once a customer has personalized after reserving via an OTA he is more likely to choose a direct channel and to continue to do so over the subsequent three stays.
```{r echo=FALSE}
t <- with(wide[wide$BOOKING_CHANNEL.1 %in% c("bookings europe b.v.", "expedia", "internet"), ],
     table(Personalization.1, BOOKING_CHANNEL.2))
knitr::kable(t, caption = "Number of Reservations by Channel on First Repeat")
knitr::kable(round(prop.table(t, margin = 1), 2), caption = "Percentage of Reservations by Channel on First Repeat")

t <- with(wide[wide$BOOKING_CHANNEL.1 %in% c("bookings europe b.v.", "expedia", "internet"), ],
          table(Personalization.1, BOOKING_CHANNEL.3))
knitr::kable(t, caption = "Number of Reservations by Channel on Second Repeat")
knitr::kable(round(prop.table(t, margin = 1), 2), caption = "Percentage of Reservations by Channel on Second Repeat")

t <- with(wide[wide$BOOKING_CHANNEL.1 %in% c("bookings europe b.v.", "expedia", "internet"), ],
          table(Personalization.1, BOOKING_CHANNEL.4))
knitr::kable(t, caption = "Number of Reservations by Channel on Third Repeat")
knitr::kable(round(prop.table(t, margin = 1), 2), caption = "Percentage of Reservations by Channel on Third Repeat")
```

